--
title: "Troubleshooting Practice"
output: pdf_document
auther: "Your Name"
date: "2-18-2020
editor_options: 
  chunk_output_type: console
---

This document contains a brief exploratory data analysis of the `flights` 
data from the `nycflights13` package that we examined earlier in the semester.
Recall this dataset contains information on all flights departing New York City
(JFK, LGA, and EWR) in 2013.

The author of this document is interested in traveling from New York City 
and used the `flights` data to plan his trip. Unfortunately, a number of coding
errors were made. Some of these errors will prevent the file from knitting,
others will fail to do what is stated in the document, and others will fail 
quietly.

Your task is to knit the document and investigate the errors. In the space
provided below, make a note of each error. When you are finished the document 
should knit completely without errors, and all code chunks should do the task 
stated in the document.

## NYC Flights Investigation

I'm planning a trip departing from one of the three New York City airports. 
Before I do any investigation, I would like to know how many flights depart 
late, how many flights arrive late, and how many flights depart late but still
manage to arrive early.

I'll use the `filter()` data-wrangling function from `dplyr` to do this.

```{r late-flights}
flights %>%
  filter(dep_delay > 0)

flights %>%
  filter(arr_delay > 0)

flights %>%
  filter(dep_delay < 0, arr_delay > 0)
```

```{r load-packages, eval = FALSE, message = FALSE}
library(tidyverse)
```


If my flight is late, I want a free meal as compensation. Some flights provide
vouchers if a flight is delayed more than three hours. How many flights 
departing New York City in 2013 led to passengers receiving meal vouchers? 

```{r late-flights}
flights %>%
  filter(dep_delay > 180)
```

I hate waiting in line at Cinnabon and when there are many flights departing 
there is a line at Cinnabon. How many flights departed during normal business
hours from 9:00 - 5:00 PM? How many departed in the evening between 5:00 PM
and 9:00 PM?

```{r}{cinnabon}
flights %>%
  filter(sched_dep_time >= 900, sched_dep_time <= 1700)

flights %>%
  filter(sched_dep_time >= 1700, sched_dep_time <= 2100)
```

I'll plan to leave in the evening to make sure I can get Cinnabon without 
waiting. Also, it looks like JFK is the only airport that has a Cinnabon so I 
will leave from there.

Now I need to decide when to leave. I'll summarize the data by month,
considering only flights departing from JFK. I'll use a `group_by()` and a
`summarize()` to find the count of flights and the median and mean of the
departure delay.

If you get an error for this code, try running **line by line** as we 
demonstrated in class to troubleshoot.

```{r jfk-delay, eval = FALSE}
jfk_monthly <- flights %>%
  filter(flights == "JFK") %>%
  group_by(month) %>%
  summarize(n = count(),
            mean(dep_delay),
            sd(dep_delay))
```

Let's plot this!

```{r plot-evening-flights}
ggplot(data = jfk_montly, 
       x = factor(month), y = mean_delay) + 
  geom_point() +
```

What percent of flights to RDU from JFK arrive on time? 

```{r rdu-ontime}
flights %>%
  filter(origin == "JFK") %>%
  filter(dest = "RDU")

flights %>%
  filter(origin == "JFK")
  filter(dest == "RDU", arr_delay <= 0)


It looks like of the 3,100 flights from JFK to RDU, 1,719 are on time, so 
about 55% of flights.

How many flights went to RDU, GSO, and CLT from the three New York City 
airports in 2013? 

```{rbarplot-by-dest}
flights %>%
  filter(dest in c("RDU", "GSO", "CLT")) %>%
  ggplot(aes(x = dest)) + 
  geom_bar()
```

I like going fast. Let's mutate a new variable for the average speed of flights
going to RDU, GSO, and CLT.

```{r boxplots-by-dest}
flights %>%
  filter(dest %in% c("RDU", "GSO", "CLT")) %>%
  mutate(mph = distance / (air_time * 60)) %>%
  ggplot(aes(x = dest, y = mph)) + 
  geom_violinplot(quantiles = 0.50)
```

Considering only flights that weren't cancelled, let's find the mean, median, 
standard deviation, and IQR of the arrival delays for each month.


``{r delays-by-month}
flights %>%
  filter(!is.na(dep_delay), !is.na(arr_delay)) %>%
group_by(month) %>%
  summarize(mean_delay = mean(arr_delay),
            median_delay = median(arr_delay),
            sd_delay = sd(arr_delay),
            IQR_delay = IQR(arr_delay)
```

The worst month to travel appears to be July. The mean delay is 16.7 
minutes, and the standard deviation is 57.1. If we consider these flights to 
be a convenience sample of the population of flights, 16.7 is an estimate of the
true population mean delay, represented by the Greek letter μ. We can represent
the true population standard deviation using $\sigma$.

