---
title: "`R` errors, troubleshooting, and getting help"
author: ""
date: ""
output: html_document
editor_options: 
  chunk_output_type: console
---

## Main Ideas

- Encountering errors and misbehaving code is common at all levels of ability.

- Employing a few simple strategies and steps to follow will help you diagnose
  and fix code problems quickly and independently.
  
- Troubleshooting code takes practice.

## Packages

```{r load-packages, message = FALSE}
library(tidyverse)
```

## Lecture Notes and Exercises

`R` or `R Markdown` will provide an error message if a problem with a
computation occurs.

Encountering errors is extremely common and figuring out how to fix them 
is frustrating and time-consuming. But errors are helpful! They provide 
you with information that will help you fix your code.

In today's lecture we will introduce a step-by-step process you can follow 
whenever you encounter trouble with code. Additionally, we will introduce a few
strategies to help you fix problematic code.

### Troubleshooting steps

**Step #1:** Don't panic! 

Errors are extremely common and fixing them quickly takes practice. Try not to 
panic or get frustrated. 

**Step #2:** Find the relevant code.

Navigate to the code chunk where the error or problem occurred. If the error 
occurred while running a code chunk this is already done. If the error occurred
while knitting, `R` usually provides a line number. If possible, navigate to the
code chunk in question so you can see both the code and error at the same time.

**Step #3:** Read the error.

Pause and read the error carefully and in full. What does it say in plain 
English? **Usually**, there is enough information provided to both diagnose and 
fix the problem.

Below are some common errors that have enough information to fix the problem.

- `...could not find function...`

`R` can't find a function that you used. Did you include a code chunk loading
the necessary packages? If you did is `eval = FALSE` included as a code chunk
option?

- `...object 'some-name-of-object' not found...`

`R` can't find `some-name-of-object`. Is `some-name-of-object` spelled and
formatted consistently, including correct capitalization? Was 
`some-name-of-object` actually created in a code chunk where `eval` is not set
to `FALSE`? Is the code chunk located **above** the code chunk where the
error occurred? Is it stored via `<-`? The "Environment" pane can be helpful 
here.

Remember, `dplyr` functions do not modify the object in place.

```{r}
mtcars_new <- mtcars %>% 
  mutate(cyl = factor(cyl))
```

- `...unexpected 'some-symbol' in...`

Here `some-symbol` can be a comma, parenthesis, bracket, etc. This means the 
code you are running is not correct syntactically. Do you have an extra comma, 
parenthesis, bracket, or other symbol? Did you make a typo?

- The `+` or `,` or ` %>% ` ends a line.

- `Error: attempt to use zero-length variable name`

This generally happens when you highlight the backticks in the code chunk 
in addition to the code. It can also occur if there is a stray `%>%` or `+` at 
the end of a code chunk.

```{r}
mtcars %>% 
  ggplot(mapping = aes(x = mpg, y = hp)) +
  geom_point() +
  labs(x = "Miles per gallon", y = '"Horse" power')
```

- `...requires the following missing aesthetics...`

`ggplot()` is missing a necessary aesthetic. Is the `ggplot()` formatted 
correctly? Is the aesthetic provided consistent with what is required? 

```{r, error=TRUE}
mtcars %>% 
  ggplot(aes(x = mpg)) +
  geom_point()
```

We are missing the y aesthetic that is required for `geom_point()`.

- `object of type 'closure' is not subsettable`

The above error is included because it is pretty common. It (usually) means
that you tried to subset a function.

Check the type of R objects:

```{r}
typeof(c("a"))
typeof(1:10)
typeof(mean)
```

```{r error=TRUE}
letters[10]
mean[2]
```

- `non-numeric argument to binary operator`

Occurs when you mix different data types in a calculation.

Often, closely reading the error will allow you to fix the problem. 

**Step #4:** Run the code line by line.

For a code chunk with a number of lines, it is sometimes tricky to tell where
exactly the error is occurring. In this situation, it is helpful to run the 
code line-by-line to figure out where the error is occurring.

The code chunk below has a few small errors. Let's demonstrate how running the
code line-by-line helps us troubleshoot.

```{r line-by-line, eval = FALSE}
mpg %>% 
  filter(class == "subcompact") %>%
  group_by(drv) %>% 
  summarize(median_cty_mpg = median(cty),
            sd_cty_mpg     = sd(cty),
            avg_cty_mpg    = mean(cty))
```

**Step #5: Examine the Documentation**

If an error is from a particular function or argument, pulling up the 
documentation is a good step.

Documentation in `R` is extremely helpful. If you want to understand what a 
function does, its arguments, or examples of usage, examine the documentation.
In many situations, it should be higher than the fifth troubleshooting step.

Documentation can be examined using `?` or `help()`. 

- Description: a general description of what the function does
- Usage: the arguments of the function and their defaults
- Arguments: a description of each argument and what it does
- Details: details about the function
- Examples: examples of function usage

Take a look at `pull()`

```{r}
mtcars %>% pull(1)
mtcars %>% pull(cyl)
```

**Step 6: Google**

Generally you are not the first person to encounter a particular error in `R`. 
A well-thought out Google search can lead to others who have encountered the
same or a similar problem and potentially a solution.

**Include** general search terms related to the error. Include quotation
marks to search for an exact phrase and include aspects of the error that are
unique. If the error is with a function in a particular package (`rvest`, 
`dplyr`, etc) include that with your search. Include `R` as a search term.

**Avoid** search terms that are specific to your current project. This includes
datasets, variables, and aspects of your personal system (file paths, etc). You
can exclude search terms using a "-" in Google.

You can also use advanced search operators. A helpful official Google link is
[here](https://support.google.com/websearch/answer/2466433?hl=en) and an 
unofficial blog post is 
[here](https://www.lifewire.com/advanced-google-search-3482174).
Check out partial search, domain search, and words by proximity.

[StackOverflow](https://stackoverflow.com/) is an extremely helpful site.
Check out the [`R` related topics](https://stackoverflow.com/questions/tagged/r)
with the tag `R`.

**Examine a Vignette**

For a broad overview of the capabilities of a package it is helpful to examine
a vignette. Vignettes are "discursive documents meant to illustrate and explain
facilities in [a] package" (R-Project).

Use `browseVignettes()` to see vignettes from all installed packages and
`browseVignettes(package = some-package)` to see vignettes from a particular
package. Use `vignette("vignette-name")` to see a particular vignette.

Let's try examining a vignette from Wednesday's lecture.

**Step #7:** Post to Slack in the correct channel

First (if possible) push your most recent work to GitHub.

In your Slack posting, state the following:

- What you are trying to accomplish.
- What issue you encountered.
- A (brief) overview of the steps you have taken to solve the issue.

Include both the code and error in a code chunk using the code formatting
option available in Slack. Enter three backticks and paste your code in
the code/markdown box.

If possible, include a reproducible example of the error.

### Reproducible examples

A reproducible example (reprex) is a very small, toy example used to recreate
the issue for yourself and others. Often, devising a reprex helps you diagnose
the issue. 

Check out [How to make a great R ReproducibleExample](https://stackoverflow.com/questions/5963269/how-to-make-a-great-r-reproducible-example)
and [How to create a Minimal, Reproducible Example](https://stackoverflow.com/help/minimal-reproducible-example).

A reprex should be:

- minimal: use as little code as possible to reproduce the issue
- complete: should include all aspects necessary to reproduce the problem
- reproducible: should generate the issue in question

Check out the examples [here](https://stackoverflow.com/questions/24576515/relative-frequencies-proportions-with-dplyr/24576703#24576703) and [here](https://stackoverflow.com/questions/35090883/remove-all-of-x-axis-labels-in-ggplot).

The function `tribble()` is helpful for creating small, toy datasets row-by-row.

Let's now get some practice with troubleshooting errors!

## Additional Resources

- [R Project Help](https://www.r-project.org/help.html)
- [Advanced Debugging](https://adv-r.hadley.nz/debugging.html)
- [Reprex package](https://www.tidyverse.org/help/#reprex-pkg)
- [How to write a reproducible example](http://adv-r.had.co.nz/Reproducibility.html)
